---
title: "RNA-seq Analysis Revision"
author: "HHN"
date: "2024-08-09"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install R libraries for Go-term visualization

```{r Libraries, include=FALSE, echo=FALSE}
library("ComplexUpset")
library("ggpubr")
library("stringr")
library("r2r")
library("gt")
library("gtExtras")
library("ComplexUpset")
library("tibble")
library("tidyr")
```

```{r function, include=FALSE, echo=FALSE}
upset_plot_reorder = function(data_list, line_treatment, line_treatment_list){
for (gene_id in data_list$target) {
  for (lines in line_treatment) {
    if(lines %in% data_list[data_list$target == gene_id, "contrast"]){
      line_treatment_list[[lines]] = append(line_treatment_list[[lines]],1)
    }else{
      line_treatment_list[[lines]] = append(line_treatment_list[[lines]],0)
    }
  }
}
  return(line_treatment_list)
}

outersect <- function(x, y) {
  sort(c(setdiff(x, y),
         setdiff(y, x)))
}
```


# Go Clusters of downregulated Genes
## Read Files

```{r Read files}

# Downregulated Clusters: 1, 2, 4, 10
cluster_01 = read.csv2(file = "./Data/GO for clusters/Cluster1_gProfiler_athaliana_07-01-2025_15-14-23__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)
cluster_02 = read.csv2(file = "./Data/GO for clusters/Cluster2_gProfiler_athaliana_07-01-2025_15-19-19__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)
cluster_04 = read.csv2(file = "./Data/GO for clusters/Cluster4_gProfiler_athaliana_07-01-2025_15-21-37__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)
cluster_10 = read.csv2(file = "./Data/GO for clusters/Cluster10_gProfiler_athaliana_07-01-2025_15-24-58__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)

cluster_1_terms = c("membrane", "transmembrane transporter activity", "lipase activity", "positive regulation of biological process", 
                    "response to hormone", "oxoacid metabolic process", "response to stress", "localization",
                    "cellular catabolic process", "endoplasmic reticulum to cytosol auxin transport", "response to abscisic acid",
                    "oligosaccharide metabolic process", "sucrose metabolic process", "glycosinolate metabolic process", "response to stimulus")

cluster_2_terms = c("plant-type vacuole", "DNA-binding transcription factor activity", "transporter activity", "MAP kinase activity",
                    "regulation of RNA biosynthetic process", "biological regulation", "response to endogenous stimulus","protein ubiquitination",
                    "rhythmic process", "carbohydrate metabolic process", "regulation of plant organ morphogenesis",
                    "response to auxin", "xyloglucan metabolic process", "response to gibberellin",
                    "negative regulation of hormone metabolic process")

cluster_4_terms = c("plasma membrane","NAD(P)H dehydrogenase complex (plastoquinone)", "DNA-binding transcription factor activity", 
                    "transcription regulator activity", "regulation of cellular process", "regulation of metabolic process", "aromatic compound biosynthetic process", "NAD(P)H dehydrogenase complex assembly", "chloroplast organization", "carbohydrate metabolic process", "response to blue light", "photomorphogenesis", "monopolar cell growth", "chloroplast RNA modification", "inorganic anion transmembrane transport")
  
cluster_10_terms = c("chloroplast", "plasma membrane", "photosystem II", "catalytic activity", "electron transporter, transferring electrons within the cyclic electron transport pathway of photosynthesis activity", "cellular process", "metabolic process", "photosynthesis", "leaf development", "carbohydrate catabolic process", "polysaccharide catabolic process", "regulation of lateral root development", "plastid localization","carbon fixation", "carbohydrate biosynthetic process")

# Filter for selected terms and reorder them
cluster_01 = cluster_01[cluster_01$term_name %in% cluster_1_terms,]
cluster_01$term_name = factor(cluster_01$term_name, levels = cluster_01$term_name[order(cluster_01$negative_log10_of_adjusted_p_value)])
cluster_01$term_ratio = paste(cluster_01$intersection_size, cluster_01$term_size, sep = "/")

# Filter for selected terms and reorder them
cluster_02 = cluster_02[cluster_02$term_name %in% cluster_2_terms,]
cluster_02$term_name = factor(cluster_02$term_name, levels = cluster_02$term_name[order(cluster_02$negative_log10_of_adjusted_p_value)])
cluster_02$term_ratio = paste(cluster_02$intersection_size, cluster_02$term_size, sep = "/")

#cluster_9_terms[!(cluster_9_terms %in% cluster_09$term_name)]

# Filter for selected terms and reorder them
cluster_04 = cluster_04[cluster_04$term_name %in% cluster_4_terms,]
cluster_04$term_name = factor(cluster_04$term_name, levels = cluster_04$term_name[order(cluster_04$negative_log10_of_adjusted_p_value)])
cluster_04$term_ratio = paste(cluster_04$intersection_size, cluster_04$term_size, sep = "/")

# Filter for selected terms and reorder them
cluster_10 = cluster_10[cluster_10$term_name %in% cluster_10_terms,]
cluster_10$term_name[cluster_10$term_name == "electron transporter, transferring electrons within the cyclic electron transport pathway of photosynthesis activity"] = "electron transporter, transferring electrons within cyclic electron transport pathway..."
cluster_10$term_name = factor(cluster_10$term_name, levels = cluster_10$term_name[order(cluster_10$negative_log10_of_adjusted_p_value)])
cluster_10$term_ratio = paste(cluster_10$intersection_size, cluster_10$term_size, sep = "/")



```

```{r Go_cluster_gene_dependancy}

dlab_theme = theme(text = element_text(size=5), axis.text = element_text(size=5),
        axis.text.x = element_text(size = 5, face = "bold", colour = "black"),
        axis.text.y = element_text(color = "black", angle = 0, hjust = 1),
        #axis.line = element_line(size = 1, linetype = "solid"),
        axis.ticks = element_line(size = 1),
        axis.title.y = element_blank(),
        legend.position="none",
        legend.text= element_text(size = 5),
        #legend.title = element_text(size = 16, face = "bold"),
        panel.background = element_blank(),
        #panel.spacing.x = unit(1, "lines"),
        #panel.spacing.y=unit(0,"lines"),
        #panel.grid =  element_blank(),
        panel.grid.major.x =  element_line(colour = "#999999", linetype = "solid"),
        panel.grid.minor.x =  element_line(colour = "#CCCCCC", linetype = "solid"),
        panel.grid.major.y =  element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.background =element_rect(fill="#6495ED"),
        plot.margin = margin(10,10,0,0),
        strip.text.x = element_text(size = 5, face = "bold", hjust = 0.5))


cluster1_plot = ggplot(cluster_01, mapping = aes(x = term_name, y = negative_log10_of_adjusted_p_value)) +
  geom_bar(aes(fill = source), stat = "identity") +
  geom_text(aes(label = term_ratio, y = negative_log10_of_adjusted_p_value + 1.5), size = 1.5) + 
  ylab("-Log10(FDR)") +
  dlab_theme +
  coord_flip() +
  facet_grid(rows = vars(source), scales = 'free_y', space = "free_y", drop = TRUE) +
  scale_x_discrete(breaks = cluster_01$term_name, labels = toupper(str_wrap(as.character(cluster_01$term_name), width = 25))) +
  scale_y_continuous(expand = c(0,0), limits = c(0,15))

cluster2_plot = ggplot(cluster_02, mapping = aes(x = term_name, y = negative_log10_of_adjusted_p_value)) +
  geom_bar(aes(fill = source), stat = "identity") +
  geom_text(aes(label = term_ratio, y =  negative_log10_of_adjusted_p_value+ 1.5), size = 1.5) + 
  ylab("-Log10(FDR)") +
  dlab_theme +
  coord_flip() +
  facet_grid(rows = vars(source), scales = 'free_y', space = "free_y", drop = TRUE) +
  scale_x_discrete(breaks = cluster_02$term_name, labels = toupper(str_wrap(as.character(cluster_02$term_name), width = 25)))+
  scale_y_continuous(expand = c(0,0), limits = c(0,15))

cluster4_plot = ggplot(cluster_04, mapping = aes(x = term_name, y = negative_log10_of_adjusted_p_value)) +
  geom_bar(aes(fill = source), stat = "identity") +
  geom_text(aes(label = term_ratio, y = negative_log10_of_adjusted_p_value+ 1.5), size = 1.5) + 
  ylab("-Log10(FDR)") +
  dlab_theme +
  coord_flip() +
  facet_grid(rows = vars(source), scales = 'free_y', space = "free_y", drop = TRUE) +
  scale_x_discrete(breaks = cluster_04$term_name, labels = toupper(str_wrap(as.character(cluster_04$term_name), width = 25)))+
  scale_y_continuous(expand = c(0,0), limits = c(0,15))

cluster10_plot = ggplot(cluster_10, mapping = aes(x = term_name, y = negative_log10_of_adjusted_p_value)) +
  geom_bar(aes(fill = source), stat = "identity") +
  geom_text(aes(label = term_ratio, y = negative_log10_of_adjusted_p_value+ 2.5), size = 1.5) + 
  ylab("-Log10(FDR)") +
  dlab_theme +
  coord_flip() +
  facet_grid(rows = vars(source), scales = 'free_y', space = "free_y", drop = TRUE) +
  scale_x_discrete(breaks = cluster_10$term_name, labels = toupper(str_wrap(as.character(cluster_10$term_name), width = 25)))+
  scale_y_continuous(expand = c(0,0), limits = c(0,60))

ggsave("cluster1.pdf", plot = cluster1_plot , units = "in", width = 4,
  height = 3.25)
ggsave("cluster2.pdf", plot = cluster2_plot, units = "in", width = 4,
  height = 3.25)
ggsave("cluster4.pdf", plot = cluster4_plot , units = "in", width = 4,
  height = 3.25)
ggsave("cluster10.pdf", plot = cluster10_plot , units = "in", width = 4,
  height = 3.25)

Down_Cluster = ggarrange(cluster1_plot, cluster2_plot, cluster4_plot, cluster10_plot, labels = c("Cluster 1", "Cluster 2", "Cluster 4", "Cluster 10"), hjust = -0.1, vjust = 1)
ggsave("Down_Cluster.pdf", plot = Down_Cluster , units = "in", width = 8, height = 6.5)
```


## Upset Plot

This is an Upset plot for differentiating PTI, ETI, PTI + ETI specific and overlapping genes.

```{r Upset Plot, echo=TRUE}
gene_list = read.csv2("./Data/Immune_Pathways/3drnaseq/X2024.08.05.11.24.04.j284/result/Significant DE genes list and statistics.csv", sep = ",", dec = ".", header = TRUE)
gene_list = gene_list[gene_list$contrast %in% c("SETI_WT.E-SETI_WT.mock", "SETI_WT.P-SETI_WT.mock", "SETI_WT.PE-SETI_WT.mock"),]

upregulated = gene_list[gene_list$up.down == "up-regulated", ]
downregulated = gene_list[gene_list$up.down == "down-regulated", ]

line_treatment = c("SETI_WT.E-SETI_WT.mock", "SETI_WT.P-SETI_WT.mock", "SETI_WT.PE-SETI_WT.mock")
line_treatment_list = list("SETI_WT.E-SETI_WT.mock" = NULL, "SETI_WT.P-SETI_WT.mock" =  NULL, "SETI_WT.PE-SETI_WT.mock" =  NULL)

upregulated_list = upset_plot_reorder(upregulated, line_treatment, line_treatment_list)
upset_plot_up = data.frame(genes = upregulated$target, ETI = upregulated_list[["SETI_WT.E-SETI_WT.mock"]], PTI = upregulated_list[["SETI_WT.P-SETI_WT.mock"]], PTI_plus_ETI = upregulated_list[["SETI_WT.PE-SETI_WT.mock"]])

upset_plot_up = upset_plot_up[!duplicated(upset_plot_up$genes),]
upset(upset_plot_up, colnames(upset_plot_up)[1:4])

var.lab3 = c(
  "PTI" = "PTI",
  "ETI" = "ETI",
  "PTI_plus_ETI" = "PTI + ETI")[sort(1:3, decreasing = TRUE)]

vlabeller <- function (variable, value) {
  return(var.lab3[value])
}

upset_plot_up = upset_plot_up[,c(1,3,2,4)]

upset_up_plot = upset(upset_plot_up, colnames(upset_plot_up[,c(4,3,2)]), name='Lines', width_ratio=0.1, labeller= vlabeller, sort_sets=FALSE, set_sizes=FALSE, themes=upset_default_themes(text=element_text(size=8))) + ggtitle('Intersection of upregulated DEGs')

# ETI specific genes
#write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 1 & upset_plot_up$PTI == 0 & upset_plot_up$PTI_plus_ETI == 0, "genes"]),
#           file = "ETI_Specific_Genes.csv")

#write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 0 & upset_plot_up$PTI == 1 & upset_plot_up$PTI_plus_ETI == 0, "genes"]),
#           file = "PTI_Specific_Genes.csv")

#write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 0 & upset_plot_up$PTI == 1 & upset_plot_up$PTI_plus_ETI == 1, "genes"]),
#           file = "PTI_plus_ETIPTI_Specific_Genes.csv")

write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 1 & upset_plot_up$PTI == 1 & upset_plot_up$PTI_plus_ETI == 1, "genes"]),
          file = "PTI_plus_ETI_plus_ETIPTI_Specific_Genes.csv")

write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 1 & upset_plot_up$PTI == 1 & upset_plot_up$PTI_plus_ETI == 0, "genes"]),
          file = "PTI_plus_ETI_Specific_Genes.csv")

write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 0 & upset_plot_up$PTI == 0 & upset_plot_up$PTI_plus_ETI == 1, "genes"]),
          file = "PTI_ETI_Specific_Genes.csv")

write.csv2(data.frame(Genes = upset_plot_up[upset_plot_up$ETI == 1 & upset_plot_up$PTI == 0 & upset_plot_up$PTI_plus_ETI == 1, "genes"]),
          file = "ETI_plus_PTIETI_Specific_Genes.csv")



ETI = read.csv2("Figures_Tables/ETI_Specific_Genes.csv")
PTI = read.csv2("Figures_Tables/PTI_Specific_Genes.csv")
PTI_plus_ETIPTI_Specific = read.csv2("Figures_Tables/PTI_plus_ETIPTI_Specific_Genes.csv")
Upset_plot_gene_list = gene_list
gene_list[gene_list %in% c(ETI, PTI, PTI_plus_ETIPTI_Specific)]

Genes = length(upset_plot_up[upset_plot_up$ETI == 0 & upset_plot_up$PTI == 0 & upset_plot_up$PTI_plus_ETI == 1, "genes"])

# pdf(file="upset_plot_up.pdf",
#     width = 4,
#   height = 3.25)
# upset_up_plot
# dev.off()

```


# Figure 5
## Read Files

```{r Read files}
gene_list = read.csv2(file = "./Data/Downregulated_ETI/Significant DE genes list and statistics.csv", sep = ",", dec = ".",
                      header = TRUE)
gene_list = gene_list[gene_list$contrast != "SETI_p.E-SETI_s.E" & gene_list$contrast != "SETI_p.E-SETI_WT.E"
          & gene_list$contrast != "SETI_s.E-SETI_WT.E", ]
gene_list$contrast = factor(gene_list$contrast)

full_gene_list = read.csv("./Data/Downregulated_ETI/Full_geneslist_logF2.csv", sep = ";", dec = ".", header = TRUE, row.names = 1)


```


## Upset Plot Downregulated Genes

Same dataset as figure 5 but with downregulated genes differentiating PAD4, SAG101 and EDS1 dependent ETI genes

```{r Upset Plot (Downregulated), echo=TRUE}

upregulated = gene_list[gene_list$up.down == "up-regulated", ]
downregulated = gene_list[gene_list$up.down == "down-regulated", ]

line_treatment = c("SETI_WT.E-SETI_WT.mock", "SETI_p.E-SETI_p.mock", "SETI_s.E-SETI_s.mock", "SETI_ps.E-SETI_ps.mock", "SETI_e.E-SETI_e.mock")
line_treatment_list = list("SETI_WT.E-SETI_WT.mock" = NULL, "SETI_p.E-SETI_p.mock" =  NULL, "SETI_s.E-SETI_s.mock" =  NULL, "SETI_ps.E-SETI_ps.mock" =  NULL,  "SETI_e.E-SETI_e.mock" =  NULL)

downregulated_list = upset_plot_reorder(downregulated, line_treatment, line_treatment_list)
upset_plot_down = data.frame(genes = downregulated$target, SETI = downregulated_list[["SETI_WT.E-SETI_WT.mock"]], PAD4 = downregulated_list[["SETI_p.E-SETI_p.mock"]], SAG101 = downregulated_list[["SETI_s.E-SETI_s.mock"]], PAD4SAG101 = downregulated_list[["SETI_ps.E-SETI_ps.mock"]],
                           EDS1 = downregulated_list[["SETI_e.E-SETI_e.mock"]])

upset_plot_down = upset_plot_down[!duplicated(upset_plot_down$genes),]
write.csv2(downregulated[downregulated$contrast == "SETI_WT.E-SETI_WT.mock",], file = "SETI_downregulated")

upset(upset_plot_down, colnames(upset_plot_down)[2:6])

var.lab3 = c(
  "SETI" = bquote(bold("SETI wt")),
  "PAD4" = expression(bold("SETI") ~ bolditalic("pad4")),
  "SAG101" = expression(bold("SETI") ~ bolditalic("sag101")), 
  "PAD4SAG101" = expression(bold("SETI") ~ bolditalic("pad4 sag101")),
  "EDS1" = expression(bold("SETI") ~ bolditalic("eds1")))[sort(1:5, decreasing = TRUE)]


vlabeller <- function (variable, value) {
  return(var.lab3[value])
}

upset_plot_down = upset_plot_down[,c(1, sort(2:6, decreasing = TRUE))]

upset_up_plot = upset(upset_plot_down, colnames(upset_plot_down[,2:6]), name='Lines', width_ratio=0.1, labeller= vlabeller, sort_sets=FALSE, set_sizes=FALSE, themes=upset_default_themes(text=element_text(size=8))) + ggtitle('Intersection of downregulated DEGs')

Genes = length(upset_plot_down[upset_plot_down$PAD4 == 0 & upset_plot_down$SAG101 == 1 & upset_plot_down$SETI == 0 & upset_plot_down$EDS1 == 0
                               & upset_plot_down$PAD4SAG101 == 0, "genes"])

# pdf(file="upset_plot_down.pdf",
#     width = 4,
#   height = 3.25)
# upset_up_plot
# dev.off()

```

# Representative Genes

```{r Read_Clusters, echo=TRUE}

# Downregulated Clusters: 1, 2, 4, 10
cluster_01 = read.csv2(file = "./Data/GO for clusters/Cluster1_gProfiler_athaliana_07-01-2025_15-14-23__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)
cluster_02 = read.csv2(file = "./Data/GO for clusters/Cluster2_gProfiler_athaliana_07-01-2025_15-19-19__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)
cluster_04 = read.csv2(file = "./Data/GO for clusters/Cluster4_gProfiler_athaliana_07-01-2025_15-21-37__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)
cluster_10 = read.csv2(file = "./Data/GO for clusters/Cluster10_gProfiler_athaliana_07-01-2025_15-24-58__intersections.csv", sep = ",", dec = ".",
                      header = TRUE)

string_transpose = function(cluster, term){
  representative = str_split_1(cluster$intersections[cluster$term_name == term], ",")
  representative = setNames(representative, rep(term, length(representative)))
  return(representative)
}

representative1 = str_split_1(cluster_01$intersections[cluster_01$term_name == "response to abscisic acid"], ",")
representative1 = setNames(representative1, rep("response to abscisic acid", length(representative1)))

representative4 = string_transpose(cluster_02, "response to gibberellin")
representative5 = string_transpose(cluster_02, "carbohydrate metabolic process")
representative6 = string_transpose(cluster_02, "regulation of plant organ morphogenesis")
representative3 = string_transpose(cluster_02, "response to auxin")

representative2 = str_split_1(cluster_04$intersections[cluster_04$term_name == "photomorphogenesis"], ",")
representative2 = setNames(representative2, rep("photomorphogenesis", length(representative2)))

representative7 = string_transpose(cluster_10, "photosynthesis")
representative8 = string_transpose(cluster_10, "leaf development")
representative9 = string_transpose(cluster_10, "regulation of lateral root development")



gene_id = c(representative1, representative2, representative3, representative4, representative5, representative6, representative7, representative8, representative9)
gene_terms = c(names(representative1), names(representative2), names(representative3), names(representative4), names(representative5), names(representative6), names(representative7), names(representative8), names(representative9))

representative_genes = data.frame(Gene_ID = gene_id , term_name = gene_terms)
representative_genes = cbind(representative_genes, full_gene_list[match(representative_genes$Gene_ID, row.names(full_gene_list)), c("SETI_WT.E.SETI_WT.mock", "SETI_p.E.SETI_p.mock", "SETI_s.E.SETI_s.mock", "SETI_ps.E.SETI_ps.mock")])
colnames(representative_genes)[3:6] = c("SETI", "PAD4", "SAG101", "PAD4_SAG101")
fig_table = representative_genes[,c("Gene_ID", "SETI", "PAD4", "SAG101", "PAD4_SAG101")]
fig_table = fig_table[!duplicated(fig_table$Gene_ID),]
#fig_table = fig_table[((fig_table$SETI < fig_table$PAD4) & (fig_table$SAG101 < fig_table$PAD4)) |
#                        ((fig_table$SETI > fig_table$PAD4) & (fig_table$SAG101 > fig_table$PAD4)),]
fig_table = fig_table[((fig_table$SETI < fig_table$PAD4) & (fig_table$SAG101 < fig_table$PAD4)),]

fig_table$SETI = round(fig_table$SETI, 3)
fig_table$PAD4 = round(fig_table$PAD4, 3)
fig_table$SAG101 = round(fig_table$SAG101, 3)
fig_table$PAD4_SAG101 = round(fig_table$PAD4_SAG101, 3)
fig_table = fig_table[order(fig_table$PAD4),]

# Remove some genes and reduce list size
fig_table1 = na.omit(fig_table[fig_table$Gene_ID %in% c(representative2,representative7,representative8,representative4[1:11], representative9),])
group2 = c(representative1[-13],representative5[1:12],representative3,representative6) #[c(representative1,representative5,representative3,representative6)]
fig_table1_1 = na.omit(fig_table[fig_table$Gene_ID %in% group2,])


```

```{r Table Figure, echo=TRUE}

table_1 = gt(fig_table1, rowname_col = "Gene_ID")|> 
  tab_row_group(
    label = "Response to gibberellin",
    rows = representative4[1:11][representative4[1:11] %in% fig_table$Gene_ID]
  ) |>
  tab_row_group(
    label = "Photosynthesis",
    rows = representative7[representative7 %in% fig_table$Gene_ID]
  ) |>
  tab_row_group(
    label = "Leaf development",
    rows = representative8[representative8 %in% fig_table$Gene_ID]
  ) |>
  tab_row_group(
    label = "Photomorphogenesis",
    rows = representative2[representative2 %in% fig_table$Gene_ID]
  ) |>
    tab_row_group(
    label = "Regulation of lateral root development",
    rows = representative9[representative9 %in% fig_table$Gene_ID]
  ) |>
  tab_stubhead(label = "Gene ID")  |>
  cols_label(
    Gene_ID = "Gene ID",
    SETI = md("SETI"),
    SAG101 = md("SETI *sag101*"),
    PAD4 = md("SETI *pad4*"),
    PAD4_SAG101 = md("SETI *pad4 sag101*")
  ) |>
  tab_style(
    style = cell_fill(color = "#d4d2d2"),  # Set the background color to grey
    locations = list(
      cells_column_labels(),
      cells_row_groups(),
      cells_stub(),
      cells_stubhead(),
      cells_column_spanners()
    )     # Target the row groups
  )

table_1_1 = gt(fig_table1_1, rowname_col = "Gene_ID")|> 
  tab_row_group( 
    label = "Response to abscisic acid",
    rows = representative1[-13][representative1[-13] %in% fig_table$Gene_ID],
    ) |>
  tab_row_group(
    label = "Response to auxin",
    rows = representative3[(representative3 %in% fig_table$Gene_ID)] 
  ) |>  
  tab_row_group(
    label = "Carbohydrate metabolic process",
    rows = representative5[1:12][representative5[1:12] %in% fig_table$Gene_ID]
  ) |>
  tab_row_group(
    label = "Regulation of plant organ morphogenesis",
    rows = representative6[representative6 %in% fig_table$Gene_ID]
  ) |>
  tab_stubhead(label = "Gene ID")  |>
  cols_label(
    Gene_ID = "Gene ID",
    SETI = md("SETI"),
    SAG101 = md("SETI *sag101*"),
    PAD4 = md("SETI *pad4*"),
    PAD4_SAG101 = md("SETI *pad4 sag101*")
  ) |>
  tab_style(
    style = cell_fill(color = "#d4d2d2"),  # Set the background color to grey
    locations = list(
      cells_column_labels(),
      cells_row_groups(),
      cells_stub(),
      cells_stubhead(),
      cells_column_spanners()
    )     # Target the row groups
  )


table_1 = tab_style(table_1,
    style = cell_text(align = "center"),
    locations = list(
      cells_body(),
      cells_row_groups(),
      cells_stub(),
      cells_column_spanners()
    )
  )

table_1 = tab_style(table_1,
    style = cell_text(align = "center", v_align = "middle"),
    locations = cells_column_labels() 
  ) |>
  tab_style(
    style = cell_text(align = "center", v_align = "middle"),
    locations = cells_stubhead()
  ) |>
  tab_style(
    style = cell_borders(
      sides = "all", # Add borders to all sides
      color = "black", # Border color
      weight = px(1)   # Border thickness
    ),
    locations = list(
      cells_body(),
      cells_row_groups(),
      cells_stub(),
      cells_column_spanners(),
      cells_source_notes(),
      cells_footnotes(),
      cells_title()
    )
  ) |>
  cols_width(
    everything() ~ px(60)  # Set all columns to 100 pixels wide
  ) |> 
  gt::tab_options(
    column_labels.border.bottom.width = px(1), # Fix column label and spanner bottom border
    column_labels.border.top.width = px(1), # Fix column label top border
    table.border.bottom.width = px(1), # Fix table bottom border
    table.border.top.width = px(1), # Fix table top border
    table_body.border.top.width = px(1), # Fix body top border
    table_body.border.bottom.width = px(1) # Fix body bottom border
  ) |>
  tab_style(
    style = cell_borders(sides = "all", color = "black", weight = px(1)),
    locations = list(cells_column_labels(everything()),
                     cells_stubhead())
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_column_labels(),
                     cells_stubhead(),
                     cells_row_groups(),
                     cells_body(columns = "Gene_ID"),
                     cells_stub())
  ) %>%
  tab_style(
    style = cell_text(size = px(8)),
    locations = list(cells_body(), cells_column_labels(), cells_stub(), cells_stubhead(), cells_row_groups()) 
  )


table_1_1 = tab_style(table_1_1,
    style = cell_text(align = "center"),
    locations = list(
      cells_body(),
      cells_row_groups(),
      cells_stub(),
      cells_column_spanners()
    )
  )

table_1_1 = tab_style(table_1_1,
    style = cell_text(align = "center", v_align = "middle"),
    locations = cells_column_labels() 
  ) |>
  tab_style(
    style = cell_text(align = "center", v_align = "middle"),
    locations = cells_stubhead()
  ) |>
  tab_style(
    style = cell_borders(
      sides = "all", # Add borders to all sides
      color = "black", # Border color
      weight = px(1)   # Border thickness
    ),
    locations = list(
      cells_body(),
      cells_row_groups(),
      cells_stub(),
      cells_column_spanners(),
      cells_source_notes(),
      cells_footnotes(),
      cells_title()
    )
  ) |>
  cols_width(
    everything() ~ px(60)  # Set all columns to 100 pixels wide
  )  |> 
  gt::tab_options(
    column_labels.border.bottom.width = px(1), # Fix column label and spanner bottom border
    column_labels.border.top.width = px(1), # Fix column label top border
    table.border.bottom.width = px(1), # Fix table bottom border
    table.border.top.width = px(1), # Fix table top border
    table_body.border.top.width = px(1), # Fix body top border
    table_body.border.bottom.width = px(1) # Fix body bottom border
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "black", style = "solid", weight = px(1)),
    locations = list(cells_column_labels(everything()),
                     cells_stubhead())
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_column_labels(),
                     cells_stubhead(),
                     cells_row_groups(),
                     cells_body(columns = "Gene_ID"),
                     cells_stub())
  ) %>%
  tab_style(
    style = cell_text(size = px(8)),
    locations = list(cells_body(), cells_column_labels(), cells_stub(), cells_stubhead(), cells_row_groups()) 
  ) %>%
  tab_options(
    table.margin.right = px(20) # Adds 20px of space to the right side
  )

    
    
  

colfunc = colorRampPalette(c("darkblue","#5c6cfa", "#d5d7ed", "white", "#f5bd73", "#f5a573", "red"))

table_1 = data_color(table_1, 
  columns = everything(),
  rows = everything(),
  method = c("bin"),
  bins = c(-5, -2, -1.5,-1, -0.9, -0.8,-0.7, -0.6, -0.5, -0.4, -0.3, -0.20, 0.20, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.5, 2, 5),
  palette = colfunc(23),
  domain = c(-5, 5))

colfunc = colorRampPalette(c("darkblue","#5c6cfa", "#d5d7ed", "white", "#f5bd73", "#f5a573", "red"))

table_1_1 = data_color(table_1_1, 
  columns = everything(),
  rows = everything(),
  method = c("bin"),
  bins = c(-5, -2, -1.5,-1, -0.9, -0.8,-0.7, -0.6, -0.5, -0.4, -0.3, -0.20, 0.20, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.5, 2, 5),
  palette = colfunc(23),
  domain = c(-5, 5))

my_tables = list(table_1, table_1_1)
gt_two_column_layout(my_tables)
gt_two_column_layout(my_tables, output = "save",
                     filename = "basic-two-col3.png", vwidth = 670, vheight = 620)


# colfunc = colorRampPalette(c("#5c6cfa","#d5d7ed", "#f5bd73", "#f5a573" ))
# table_1_2 = data_color(table_1, 
#   columns = everything(),
#   rows = everything(),
#   method = c("numeric"),
#   bins = 4,
#   quantiles = 20,
#   direction = "row",
#   palette = colfunc(4))

#gtsave(table_1_2, "table_1_2.html")
```

# Immune Pathways

# Involvement of CNGCs in PTI, ETI and PTI + ETI
```{r Table Figure, echo=TRUE}
immune_pathway_full_list =  tibble(read.csv2("Immune_pathway_genes.csv")[,c(-1,-4)])
immune_pathway_full_list$contrast = factor(immune_pathway_full_list$contrast)
immune_pathway_full_list = pivot_wider(immune_pathway_full_list, names_from = contrast, names_expand = TRUE, values_from = c(log2FC))
immune_pathway_full_list = immune_pathway_full_list[,c("target", "SETI_WT.E-SETI_WT.mock",  "SETI_WT.P-SETI_WT.mock","SETI_WT.PE-SETI_WT.mock")]
names(immune_pathway_full_list) = c("Gene_ID", "PTI", "ETI", "PTI + ETI")
cngcs_id = read.csv2("./Data/Representative_Genes/CNGCs.csv")
cngc_list = cbind(cngcs_id, immune_pathway_full_list[match(cngcs_id$Gene_ID, immune_pathway_full_list$Gene_ID),])
fig_table = cngc_list[,c("Gene_Name", "PTI", "ETI", "PTI + ETI")]
#fig_table = fig_table[order(fig_table$ETI),]
fig_table = na.omit(fig_table)
fig_table_long = pivot_longer(fig_table, !Gene_Name, names_to = "Treatment", values_to = "Log2FC")

table_1 = gt(fig_table, rowname_col = "Gene_Name")|> 
  tab_row_group( 
    label = "CNGCs",
    rows = fig_table$Gene_Name,
    ) |>
  tab_stubhead(label = "Gene ID")

table_1 = tab_style(table_1,
    style = cell_text(align = "center"),
    locations = cells_row_groups()
  )

table_1 = tab_style(table_1,
    style = cell_text(align = "center"),
    locations = cells_column_labels() 
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_stubhead()
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body()
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body()
  ) |>
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom", "left", "right"),
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_body()
  ) |>
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom", "left", "right"),
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_stub()) |>
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom", "left", "right"),
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_stub()) |>
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom", "left", "right"),
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_column_labels()) |>
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom", "left", "right"),
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_stubhead())  |>
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom", "left", "right"),
      color = "black",
      weight = px(1.5),
      style = "solid"
    ),
    locations = cells_row_groups()) 
    
  

colfunc = colorRampPalette(c("#5c6cfa", "#d5d7ed", "white", "#f5bd73", "#f5a573"))

table_1_1 = data_color(table_1, 
  columns = everything(),
  rows = everything(),
  method = c("bin"),
  bins = c(-6, -3, -0.5, 0.5, 3, 6),
  quantiles = 20,
  palette = colfunc(8),
  domain = c(-6, 6))

colfunc = colorRampPalette(c("#5c6cfa","white", "#f5a573" ))
table_1_2 = data_color(table_1, 
  columns = everything(),
  rows = everything(),
  direction = "row",
  palette = colfunc(4))

gtsave(table_1_1, "CNGC_log2F_absolute_values.png")
#gtsave(table_1_2, "CNGC_log2F_comparison.html")
```

# CNGC Barplot

```{r Barplot, echo=TRUE}
# p-value
immune_pathway_full_list =  tibble(read.csv2("Immune_pathway_genes.csv")[,c(-1)])
immune_pathway_full_list$contrast[immune_pathway_full_list$contrast == "SETI_WT.E-SETI_WT.mock"] = "ETI"
immune_pathway_full_list$contrast[immune_pathway_full_list$contrast == "SETI_WT.P-SETI_WT.mock"] = "PTI"
immune_pathway_full_list$contrast[immune_pathway_full_list$contrast == "SETI_WT.PE-SETI_WT.mock"] = "PTI + ETI"
immune_pathway_full_list = immune_pathway_full_list[immune_pathway_full_list$contrast == "PTI" | immune_pathway_full_list$contrast == "ETI" | immune_pathway_full_list$contrast == "PTI + ETI",]
immune_pathway_full_list$contrast = factor(immune_pathway_full_list$contrast)
cngcs_id = read.csv2("./Data/Representative_Genes/CNGCs.csv")
cngc_list = immune_pathway_full_list[which(immune_pathway_full_list$target %in% cngcs_id$Gene_ID),]
cngc_list$Gene_Name = NA
for (cngc_gene in cngcs_id$Gene_ID) {
  cngc_list$Gene_Name[grep(cngc_gene, cngc_list$target)] = cngcs_id$Gene_Name[cngcs_id$Gene_ID == cngc_gene]
}

dlab_theme = theme(axis.title.y = element_text(margin = margin(r = 10)), text = element_text(size=12), axis.text = element_text(size=12),
        axis.text.x = element_text(margin = margin(t = 5), size = 12, colour = "black", face="italic"),
        axis.text.y = element_text(color = "black"),
        axis.line.x = element_blank(),
        axis.ticks = element_blank(),
        #axis.ticks = element_line(size = 1),
        axis.title.x = element_blank(),
        legend.position="bottom", 
        legend.text= element_text(size = 12),
        #legend.title = element_text(size = 16, face = "bold"),
        panel.background = element_blank(),
        panel.spacing.x = unit(1, "lines"),
        panel.spacing.y= unit(0,"lines"),
        #panel.grid =  element_blank(),
        legend.title = element_blank(),
        panel.grid.minor.x =  element_blank(),
        panel.grid.major.x =  element_blank(),
        panel.grid.major.y =  element_blank(),
        panel.grid.minor.y =  element_line(colour = "#999999"),
        #panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
        strip.background =element_rect(fill="white"),
        strip.text.x = element_text(size = 12, hjust = 0.5))

cngc_list$Gene_Name = factor(cngc_list$Gene_Name, level = paste("CNGC", 1:20,sep = ""))
cngc_list$significance = ""
cngc_list$significance[cngc_list$adj.pval <= 0.01 & (cngc_list$log2FC >= 1 | cngc_list$log2FC <= -1)] = "*"

# Add missing CNGCs as NA (Prefiltered in 3D-RNA seq)
c("CNGC7", "CNGC8", "CNGC16", "CNGC18") #cngcs_id$Gene_Name[-na.omit(match(cngc_list$Gene_Name, cngcs_id$Gene_Name))]
missing_cngcs = tibble(Gene_Name = rep(c("CNGC7", "CNGC8", "CNGC16", "CNGC18"), times = 3), contrast = rep(c("PTI", "ETI", "PTI + ETI"), times = 4))
missing_cngcs$log2FC = 0
missing_cngcs$significance = "NA"
missing_cngcs$adj.pval = NA
missing_cngcs$target = NA
cngc_list = rbind(cngc_list, missing_cngcs)

cngc_barplot = ggplot(cngc_list, mapping = aes(x = Gene_Name, y = log2FC, fill = contrast)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.75) + 
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#12A27B"),
                    name = "Treatment", labels = c("PTI", "ETI", "PTI + ETI")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  geom_hline(yintercept=0) +
  geom_text(data = cngc_list[cngc_list$log2FC > 0,], mapping = aes(x = Gene_Name, y = log2FC, label = significance), position = position_dodge(width = 0.75), size = 5) +
   geom_text(data = cngc_list[cngc_list$log2FC < 0 | cngc_list$significance == "",], mapping = aes(x = Gene_Name, y = log2FC - 0.3, label = significance), position = position_dodge2(width = 0.75), size = 5) +
  geom_text(data = cngc_list[cngc_list$significance == "NA" & cngc_list$contrast == "ETI",], mapping = aes(x = Gene_Name, y = log2FC + 0.3, label = significance), size = 5) +
  dlab_theme 

ggsave("cngc_barplot.pdf", plot = cngc_barplot, units = "in", width = 10,
  height = 4.5)

```


```{r Table Legend, echo=TRUE}

intervals = c("-5 to -2","-2 to -1.5","-1.5 to -1","-1 to -0.9","-0.9 to -0.8","-0.8 to -0.7","-0.7 to -0.6","-0.6 to -0.5","-0.5 to -0.4","-0.4 to -0.3",
  "-0.3 to -0.2","-0.2 to 0.2","0.2 to 0.3","0.3 to 0.4","0.4 to 0.5","0.5 to 0.6","0.6 to 0.7","0.7 to 0.8","0.8 to 0.9","0.9 to 1",
  "1 to 1.5","1.5 to 2","2 to 5")

xl <- 1
yb <- 0
xr <- 1.1
yt <- 2

pdf("my_legend.pdf", width = 8, height = 6)
par(mar=c(5.1,0.5,4.1,0.5))
test = plot(NA,type="n",ann=FALSE,xlim=c(0.9,2),ylim=c(0,2.3),xaxt="n",yaxt="n",bty="n")
rect(
     xl,
     head(seq(yb,yt,(yt-yb)/23),-1),
     xr,
     tail(seq(yb,yt,(yt-yb)/23),-1),
     col=colfunc(23)
    )

text(
     x = (xl + xr) / 2, # X-coordinate (center of the rectangle)
     y = (head(seq(yb, yt, (yt - yb) / 23), -1) + tail(seq(yb, yt, (yt - yb) / 23), -1)) / 2, # Y-coordinate (center of the rectangle)
     labels = intervals, # Labels to add (1 to 23)
     cex = 0.7, # Text size
     col = c(rep("white", 5), rep("black", 18)) # Text color (change if needed)
    )

title(main = "Log2FC")

dev.off()

colfunc = colorRampPalette(c("darkblue","#5c6cfa", "#d5d7ed", "white", "#f5bd73", "#f5a573", "red"))

table_1_1 = data_color(table_1_1, 
  columns = everything(),
  rows = everything(),
  method = c("bin"),
  bins = c(-5, -2, -1.5,-1, -0.9, -0.8,-0.7, -0.6, -0.5, -0.4, -0.3, -0.20, 0.20, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.5, 2, 5),
  palette = colfunc(23),
  domain = c(-5, 5))


```